Verifying Cache Coherence Protocols for an Infinite Family of Cache Hierarchies

### 2014-02-06 paper/System.tex
spell in later half of paper

In addition to the permission states, each cache also maintains a
temporary \emph{wait state} for each address to deal with transitioning between
permission levels. The wait state consists of two parts: (a) a Boolean
representing whether the cache has sent a request to its parent cache earlier
and is waiting for a response (denoted by $c.\wait[a]$ for address $a$) and (b)
one of the three values \Mo, \Sh{} or \In{} representing the state upgrade that
the cache is waiting to make (denoted by $c.\waitS[a]$ for address $a$).
Similarly, each non-leaf cache has the corresponding wait states for its
directory.  $p.\dwait[c][a]$ denotes that cache $p$ is waiting for a response
from its child $c$ for address $a$, and $p.\dwaitS[c][a]$ denotes the state to
which $p$ is waiting for $c$ to downgrade. Both \wait{} and \dwait{} are
+\False{} initially for all caches and addresses,
because no cache is waiting for any response from either its child or its
parent.

@@

Note that the atomic transitions in Figure~\ref{trans} access only local states,
\ie{} a transition can read or write states only corresponding to a single cache
and/or the channel the cache is connected to. This restriction allows such a
system to be implemented directly into hardware. Bluespec SystemVerilog (BSV),
for instance, directly converts these transitions into efficient synchronous
hardware. Though these transitions logically happen one by one, BSV executes
several transitions simultaneously, as long as they do not access the same
state, \ie{} do not \emph{conflict}~\cite{Hoe:TCAD,HoeArvind:TRSSynthesis1}.
Sometimes, these transitions (both the guards and the actions) may not be
amenable to single cycle implementations in hardware. For example, writing a
register array can actually take several hardware clock cycles, though
logically it is a single transition. Karczmarek \etal~\cite{Karczmarek} has
provided a scheme to convert atomic transitions into synchronous hardware in
which each transition can span several clock cycles.

>>>>

In addition to the permission states, each cache also maintains a
temporary \emph{wait state} for each address to deal with transitioning between
permission levels. The wait state consists of two parts: (a) a Boolean
representing whether the cache has sent a request to its parent cache earlier
and is waiting for a response (denoted by $c.\wait[a]$ for address $a$) and (b)
one of the three values \Mo, \Sh{} or \In{} representing the state upgrade that
the cache is waiting to make (denoted by $c.\waitS[a]$ for address $a$).
Similarly, each non-leaf cache has the corresponding wait states for its
directory.  $p.\dwait[c][a]$ denotes that cache $p$ is waiting for a response
from its child $c$ for address $a$, and $p.\dwaitS[c][a]$ denotes the state to
which $p$ is waiting for $c$ to downgrade. Both \wait{} and \dwait{} are
\False{} initially for all caches and addresses,
because no cache is waiting for any response from either its child or its
parent.

@@

Note that the atomic transitions in Figure~\ref{trans} access only local states,
\ie{} a transition can read or write states only corresponding to a single cache
and/or the channel the cache is connected to. This restriction allows such a
system to be implemented directly into hardware. Bluespec SystemVerilog (BSV),
for instance, directly converts these transitions into efficient synchronous
hardware. Though these transitions logically happen one by one, BSV executes
several transitions simultaneously, as long as they do not access the same
state, \ie{} do not \emph{conflict}~\cite{Hoe:TCAD,HoeArvind:TRSSynthesis1}.
Sometimes, these transitions (both the guards and the actions) may not be
amenable to single cycle implementations in hardware. For example, writing a
register array can actually take several hardware clock cycles, though
logically it is a single transition. Karczmarek \etal \cite{Karczmarek} has
provided a scheme to convert atomic transitions into synchronous hardware in
which each transition can span several clock cycles.
